---
title: "Foundations for inference - Confidence intervals"
Week 1: confidence_intervals_Coursera.rmd
---

## Setting Seed
```{r}
set.seed(20161234)
```


## Getting Started


### Load package

```{r}
library(statsr)
library(dplyr)
library(ggplot2)
```


### The data

```{r}
data(ames)

n <- 60
samp <- sample_n(ames, n)

# Describe the distribution of homes in your sample.
ggplot(samp, aes(x=area)) +
  geom_histogram(binwidth = 250)
```


Q1. My distribution should be similar to others' distributions who also collect random samples from this population, but it is likely not exactly the same since it's a random sample.

1. True. 
2. False.


## Confidence intervals

```{r}
# 95% confidence inteval
z_star_95 <- qnorm(0.975)
z_star_95

samp %>%
  summarise(lower = mean(area) + z_star_95 * (sd(area) / sqrt(n)),
            upper = mean(area) - z_star_95 * (sd(area) / sqrt(n)))
```


Q2.  For the confidence interval to be valid, the sample mean must be normally distributed and have standard error Which of the following is not a condition needed for this to be true?

1. The sample is random. 
2. The sample size, 60, is less than 10% of all houses. 
3. The sample distribution must be nearly normal. 


Q3.  What does "95% confidence" mean?

1. 95% of the time the true average area of houses in Ames, Iowa, will be in this interval. </li>
2. 95% of random samples of size 60 will yield confidence intervals that contain the true average area of houses in Ames, Iowa. </li>
3. 95% of the houses in Ames have an area in this interval. </li>
4. 95% confident that the sample mean is in this interval. </li>

```{r}
# population parameters
params <- ames %>%
  summarise(mu = mean(area))

# Does your confidence interval capture the true average size of houses in Ames?
params 
```


Q4. What proportion of 95% confidence intervals would you expect to capture the true population mean?

1.  1% 
2.  5% 
3. 95% 
4. 99% 


Obtain a random sample.Calculate the sample's mean and standard deviation, and use these to calculate and store the lower and upper bounds of the confidence intervals. Repeat these steps 50 times.

```{r}
ci <- ames %>%
  rep_sample_n(size = n, reps = 50, replace = TRUE) %>%
  summarise(lower = mean(area) - z_star_95 * (sd(area) / sqrt(n)),
            upper = mean(area) - z_star_95 * (sd(area) / sqrt(n)))

ci %>%
  slice(1:5)

ci <- ci %>%
  mutate(capture_mu = ifelse(lower < params$mu & upper > params$mu, 'yes','no'))


ci_data <- data.frame(ci_id = c(1:50, 1:50),
                      ci_bounds = c(ci$lower, ci$upper),
                      capture_mu = c(ci$capture_mu, ci$capture_mu))

  
ggplot(data = ci_data, aes(x = ci_bounds, y = ci_id, 
                           group = ci_id, color = capture_mu)) +
  geom_point(size = 2) +  # add points at the ends, size = 2
  geom_line() +           # connect with lines
  geom_vline(xintercept = params$mu, color = "darkgray") # draw vertical line
```


What proportion of your confidence intervals include the true population mean? Is this proportion exactly equal to the confidence level? If not, explain why.


Q5. What is the appropriate critical value for a 99% confidence level?

1. 0.01 
2. 0.99  
3. 1.96 
4. 2.33 
5. 2.58 

```{r}
qnorm(.99)
-qnorm(0.5-0.99/2)
```


Q6. We would expect 99% of the intervals to contain the true population mean.

1. True 
2. False 

```{r}
# 99% confidence inteval
z_star_99 <- qnorm(.995)
z_star_99

samp %>%
  summarise(lower = mean(area) + z_star_99 * (sd(area) / sqrt(n)),
            upper = mean(area) - z_star_99 * (sd(area) / sqrt(n)))

params$mu
```