---
title: "Multiple linear regression"
Week 3: multiple_regression_Courera.Rmd

references:
- id: Hamermesh2005
  title: Beauty in the Classroom - Instructors' Pulchritude and Putative Pedagogical Productivity
  author:
  - family: Hamermesh
    given: Daniel S.
  - family: Parker
    given: Amy
  volume: 24
  URL: 'http://www.sciencedirect.com/science/article/pii/S0272775704001165'
  DOI: 10.1016/j.econedurev.2004.07.013
  publisher: Economics of Education Review
  ISSN: 0272-7757
  issue: 4
  page: 369-376
  type: article-journal
  issued:
    year: 2005
    month: 8
- id: Gelman2007
  title: Data Analysis Using Regression and Multilevel/Hierarchical Models
  author:
  - family: Gelman
    given: Andrew
  - family: Hill
    given: Jennifer
  publisher: Cambridge University Press
  city:
  type: book
  issued:
    year: 2007
  edition: 1
  ISBN: 9780521686891
---


## Grading the professor

Many college courses conclude by giving students the opportunity to evaluate the course and the instructor anonymously. In this lab we will analyze the data from this study in order to learn what goes into a positive professor evaluation.


## Getting Started


### Load packages

```{r load-packages, message=FALSE}
library(statsr)
library(dplyr)
library(ggplot2)
library(GGally)
```


### The data

The data were gathered from end of semester student evaluations for a large 
sample of professors from the University of Texas at Austin. 

```{r load-data, message=FALSE}
data(evals)
names(evals)
```

variable         | description
---------------- | -----------
`score`          | average professor evaluation score: (1) very unsatisfactory - (5) excellent.
`rank`           | rank of professor: teaching, tenure track, tenured.
`ethnicity`      | ethnicity of professor: not minority, minority.
`gender`         | gender of professor: female, male.
`language`       | language of school where professor received education: english or non-english.
`age`            | age of professor.
`cls_perc_eval`  | percent of students in class who completed evaluation.
`cls_did_eval`   | number of students in class who completed evaluation.
`cls_students`   | total number of students in class.
`cls_level`      | class level: lower, upper.
`cls_profs`      | number of professors teaching sections in course in sample: single, multiple.
`cls_credits`    | number of credits of class: one credit (lab, PE, etc.), multi credit.
`bty_f1lower`    | beauty rating of professor from lower level female: (1) lowest - (10) highest.
`bty_f1upper`    | beauty rating of professor from upper level female: (1) lowest - (10) highest.
`bty_f2upper`    | beauty rating of professor from second upper level female: (1) lowest - (10) highest.
`bty_m1lower`    | beauty rating of professor from lower level male: (1) lowest - (10) highest.
`bty_m1upper`    | beauty rating of professor from upper level male: (1) lowest - (10) highest.
`bty_m2upper`    | beauty rating of professor from second upper level male: (1) lowest - (10) highest.
`bty_avg`        | average beauty rating of professor.
`pic_outfit`     | outfit of professor in picture: not formal, formal.
`pic_color`      | color of professor's picture: color, black & white.


Q1. Is this an observational study or an experiment?

1) Observational study 
2) Experiment 


Q2. The original research question posed in the paper is whether beauty leads directly to the differences in course evaluations.  Given the study design, should the question be rephrased? If so, how? 

1) No, the question is worded accurately.
2) Yes, revise wording to "Is there an association between beauty and course    evaluations?" 
3) Yes, revise wording to "Does beauty score increase the professor's course    evaluations?"  
4) Yes, revise wording to "Does beauty score decrease the professor's course    evaluations?" 


## Exploring the data

Q3. Which of the following statements is **false** about the distribution of `score`? 

1) The median of the distribution is 4.3.
2) 25\% of the students gave their professors a score of over 4.6.  
3) 11 of students gave a professor a score below 3.  
4) The left skewness of the data suggests that the students are less likely to rate the professors highly. 

```{r score-dist}
summary(evals$score)
```


**Exercise:** Excluding `score`, select two other variables and describe their 
relationship using an appropriate visualization (scatterplot, side-by-side boxplots, 
or mosaic plot).

```{r two-vars-eda}
ggplot(data = evals, aes(y = score, x = rank)) +
  geom_boxplot()
```


## Simple linear regression

The fundamental phenomenon suggested by the study is that better looking teachers
are evaluated more favorably. Let's create a scatterplot to see if this appears 
to be the case:

```{r scatter-score-bty_avg}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_point()
```


**Exercise:** Replot the scatterplot, but this time replace the `geom_point()` layer 
with a `geom_jitter()` layer. (Use `?geom_jitter` to learn more.)  What was misleading 
about the initial scatterplot?

```{r jitter}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter()
```


Fit a linear model called `m_bty` to predict average professor 
score by average beauty rating and add the line to your plot using the following. If 
you do not remember how to do this, refer to the previous lab.

```{r scatter-score-bty_avg-line}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = TRUE)
```


The blue line is the model. The shaded gray area around the line tells us about the 
variability we might expect in our predictions. To turn that off, use `se = FALSE`.

```{r scatter-score-bty_avg-line-no-se}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE)
```


**Exercise:** Print a summary of the linear model, write out the equation, and 
interpret the slope. 

```{r summary-model}
# type your code for the Exercise here, and Knit
m_bty <- lm(score ~ bty_avg, data = evals)
summary(m_bty)

```


Q4. Average beauty score is a statistically significant predictor of evaluation score. 

1) True 
2) False 


Q5. Use residual plots to evaluate whether the conditions of least squares regression are reasonable.  Which of the following statements is an incorrect analysis of the residual plots and conditions? 

1) Linear association: The residuals plot shows a random scatter. 
2) Constant variance of residuals: No fan shape in residuals plot. 
3) Nearly normal residuals: Residuals are right skewed, but the sample size is large, so this may not be an important violation of    
   conditions. 
4) Independent observations: Classes sampled randomly, no order effect. 

```{r diag-score-bty_avg}
ggplot(data = m_bty, aes(x = .resid)) +
  geom_histogram(binwidth = .1) +
  xlab("Residuals")
```


## Multiple linear regression

Let's take a look at the relationship between one of these scores and the average 
beauty score.

```{r bty-rel}
ggplot(data = evals, aes(x = bty_f1lower, y = bty_avg)) +
  geom_jitter()

evals %>% 
  summarise(cor(bty_avg, bty_f1lower))
```


We can actually take a look at the relationships between all beauty variables (columns 13 through 19) using the 
following command:

```{r bty-rels, fig.width = 12, fig.height= 5}
ggpairs(evals, columns = 13:19)
```


These variables are collinear (correlated), and adding more than one of these 
variables to the model would not add much value to the model. 

In order to see if beauty is still a significant predictor of professor score 
after we've accounted for the gender of the professor, we can add the gender 
term into the model.

```{r scatter-score-bty_avg_gender}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```


Q6. P-values and parameter estimates should only be trusted if the conditions for the regression are reasonable. Using diagnostic plots, we can conclude that the conditions for this model are reasonable. 

1) True 
2) False 

```{r diag-mlr-model}
ggplot(data = m_bty_gen, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
```


**Exercise:** Print a summary of the multiple linear regression model.  Is `bty_avg` 
still a significant predictor of `score`?  Has the addition of `gender` to the 
model changed the parameter estimate for `bty_avg`?

```{r summary-mlr-model}
summary(m_bty)
summary(m_bty_gen)
```


`gender` from having the values of `female` and `male` to being an 
indicator variable called `gendermale` that takes a value of $0$ for females and
a value of $1$ for males. (Such variables are often referred to as "dummy" 
variables.)

\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (0) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg\end{aligned}
\]


Q7. For two professors (one male and one female) who received the same beauty rating, the male professor is predicted to have the higher course evaluation score than the female.

1) True 
2) False


**Exercise:** Create a new model called `m_bty_rank` with `gender` removed and 
`rank` added in.  How does R appear to handle categorical variables that have 
more than two levels? Note that the rank variable has three levels: teaching, 
tenure track, tenured.

```{r new-mlr-model}
m_bty_rank <- lm(score ~ bty_avg+rank, data = evals)
summary(m_bty_rank)
  
```


Q8. Which of the following is the correct order of the three levels of rank if we were to order them from lowest predicted course evaluation score to highest predicted course evaluation score? 

1) Teaching, Tenure Track, Tenured 
2) Tenure track, Tenured 
3) Tenure Track, Tenured, Teaching 
4) Teaching, Tenured, Tenure Track 


## Prediction

Suppose we want to use the model we created earlier, `m_bty_gen` to predict 
the evaluation score for a professor, Dr. Hypo Thetical, who is a male tenure track
professor with an average beauty of 3.

If we wanted to do this by hand, we would simply plug in these values into the
linear model.

We can also calculate the predicted value in R.

First, we need to create a new data frame for this professor.
```{r new-prof}
newprof <- data.frame(gender = "male", bty_avg = 3)
```

Note that I didn't need to add `rank = "tenure track"` to this data frame since
this variable is not used in our model.

Then, I can do the prediction using the `predict` function:
```{r new-prof-predict}
predict(m_bty_gen, newprof)
```

We can also construct a prediction interval around this prediction, which will
provide a measure of uncertainty around the prediction.
```{r new-prof-predict-interval}
predict(m_bty_gen, newprof, interval = "prediction", level = 0.95)
```

Hence, the model predicts, with 95% confidence, that a male professor with an 
average beauty score of 3 is expected to have an evaluation score between 3.1 
and 5.18. 


## The search for the best model

We will start with a full model that predicts professor score based on rank, 
ethnicity, gender, language of the university where they got their degree, age, 
proportion of students that filled out evaluations, class size, course level, 
number of professors, number of credits, average beauty rating, outfit, and 
picture color.

**Exercise:** Which variable would you expect to have the highest p-value in this model? Why? 
*Hint:* Think about which variable would you expect to not have any association 
with the professor score.


```{r m_full, tidy = FALSE}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```


Q9. Which of the following is the correct intrepetation of the coefficient associated with the ethnicity variable. 

1) 0.12 points lower than minority professors, all else held constant. 
2) 0.12 points higher than minority professors, all else held constant.  
3) 0.02 points lower than minority professors, all else held constant.
4) 0.02 points higher than minority professors, all else held constant. 


**Exercise:** Drop the variable with the highest p-value and re-fit the model. 
Did the coefficients and significance of the other explanatory variables change? 

```{r p-val-select}
m_full2 <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full2)
```


Now we try a different model selection method: adjusted $R^2$. Create a new model, 
`m1`, where you remove `rank` from the list of explanatory variables. Check out the 
adjusted $R^2$ of this new model and compare it to the adjusted $R^2$ of the full model.

```{r eval=FALSE}
m1 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg, data = evals)
summary(m1)$adj.r.squared
```

Then, try dropping the next variable from the full model (`ethnicity`):

```{r eval=FALSE}
m2 = lm(score ~ rank + gender + language + age + cls_perc_eval + 
    cls_students + cls_level + cls_profs + cls_credits + bty_avg, data = evals)
summary(m2)$adj.r.squared
```


Q10. Elimination of which variable from the full model yielded the highest adjusted R-squared? 

1) `bty_avg`
2) `cls_profs` 
3) `cls_students`
4) `rank` 

```{r}
m3 = lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits 
             + pic_outfit + pic_color, data = evals)
summary(m3)$adj.r.squared
```

```{r}
m4 = lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m4)$adj.r.squared
```

```{r}
m5 = lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m5)$adj.r.squared
```

```{r}
m6 = lm(score ~ ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m6)$adj.r.squared
```

## Variables with higher than significant level
```{r}
m7 <- lm(score ~ rank + ethnicity  
             + cls_students + cls_level + cls_profs  
             + pic_outfit, data = evals)
summary(m7)$r.squared
```

## Variables with lower than significant level
```{r}
m8 <- lm(score ~  gender + language + age + cls_perc_eval 
             + cls_credits + bty_avg 
             + pic_color, data = evals)
summary(m8)$r.squared
```