---
title: "Introduction to Inference"
week 2 :Introduction to data
---

## Getting stated

## load packages

```{r setup}
library(statsr)
library(dplyr)
# filter(), arrange(), select(), distinct(), mutate()
# summarise(), sample_n()
library(ggplot2)
```


### Data

```{r }
# Departure delays in flights to Raleigh-Durham
ggplot(data=nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 150) #geom = geometric object

# Departure delays headed to RUB only -> filter
rdu_flights <- nycflights %>% filter(dest == 'RDU')
ggplot(data = rdu_flights, aes(x = dep_delay))+
  geom_histogram()

# numerical summaries
# mean, median, sd, var, IQR, range, min, max
rdu_flights %>% summarise(mean_dd = mean(dep_delay),
                          sd_dd = sd(dep_delay), n=n())
```


## Questions 

## Q1. Create a new data frame that includes flights headed to SFO in February, and save this data frame as `sfo_feb_flights`. How many flights meet these criteria? 

```{r }
# fligts headed to San Francisco
sfo_feb_flights <- nycflights %>% filter(dest == 'SFO', month == 2)
nrow(sfo_feb_flights) 
```


## Q2. 2. Make a histogram and calculate appropriate summary statistics for **arrival** delays of `sfo_feb_flights`.

```{r}
ggplot(data = sfo_feb_flights, aes(x=arr_delay)) +
  geom_histogram(bins=100)

# statistics for various group in data frame -> group_by
rdu_flights %>%
  group_by(origin) %>%
  summarise(mean_dd = mean(dep_delay), sd_dd = sd(dep_delay), n = n())
```


## Q3.  Calculate the median and interquartile range for `arr_delay`s of flights in the `sfo_feb_flights` data frame, grouped by carrier. Which carrier is the has the hights IQR of arrival delays?

```{r}
# sf0-feb-fligts-arrival-delays-carrier
sfo_feb_flights %>% 
  group_by(carrier) %>% 
  summarise(median = median(arr_delay),IQR = IQR(arr_delay)) %>% 
  arrange(desc(IQR)) # order
```


## Q4. Which month has the highest average departure delay from an NYC airport? 

```{r}
# mean-dep-delay-months
nycflights %>%
  group_by(month) %>%
  summarise(mean_dd = mean(dep_delay)) %>%
  arrange(desc(mean_dd))
```


## Q5. Which month has the highest median departure delay from an NYC airport?

```{r}
# highest-median-dep-delay-month -> 12
nycflights %>%
  group_by(month) %>%
  summarise(median_dd = median(dep_delay)) %>%
  arrange(desc(median_dd))
```


## Q6. Is the mean or the median a more reliable measure for deciding which month(s) to avoid flying if you really dislike delayed flights, and why

```{r}
# delay-month-box
ggplot(nycflights, aes(x=factor(month), y=dep_delay))+
  geom_boxplot()
```


## Q7. time departure percentage?

```{r}
# which airport has the best on time departure rate

# step1. classify each flight as "on time" or "delayed" -> mutate
nycflights <- nycflights %>%
  mutate(dep_type = ifelse(dep_delay < 5, 'on time', 'delayed'))
          # mutate : create new variable

# step 2. group flights by origin airport,
# then calculate on time departure rates for each origin airport,
# and finally arrange the airports in descending order for on time departure percentage.
nycflights %>%
  group_by(origin) %>%
  summarise(ot_dep_rate = sum(dep_type == 'on time') / n()) %>%
  arrange(desc(ot_dep_rate))

ggplot(data = nycflights, aes(x = origin, fill = dep_type))+
  geom_bar()
```


# # Q8. What is the tail number of the plane with the fastest average speed?

```{r}
nycflights <- nycflights %>%
  mutate(avg_speed = distance / air_time) %>%
  arrange(desc(avg_speed))
select(nycflights, avg_speed, tailnum)
```


## Q9. Make a scatterplot of `avg_speed` vs. `distance`. 

```{r}
ggplot(data = nycflights, aes(y=avg_speed, x=distance))+
  geom_point()
```


# # Q10. "on time" if it gets to the destination on time or earlier than expected, regardless of any departure delays. Mutate the data frame to create a new variable called `arr_type` with levels `"on time"` and "delayed" Then, determine the on time arrival percentage based on whether the flight departed on time or not.'  What proportion of flights  that were `"delayed"` departing arrive `"on time"`?

```{r}
nycflights <- nycflights %>%
  mutate(arr_type = ifelse(arr_delay <= 0, 'on time', 'delayed'))
nycflights %>%
  group_by(dep_type) %>%
  summarise(ot_arr_rate = sum(arr_type == 'on time') / n())
```
