---
title: "Probability Distributions"
week 4 : Example Exploratory Data Analysis
---

# Goal of this tutorial document
# 1) reading in data
# 2) dealing with blanks
# 3) dealing with factors

### Data

```{r}
load('selected_nzes2011.Rdata')
library(dplyr)
```


### Analysis

## Q1. the relationship between the party the person voted for

```{r}
selected_nzes2011 %>%
  select(jpartyvote, jdiffvoting, singlefav) %>%          # object 'singlefav' not found
  str()

selected_nzes2011
names(selected_nzes2011)                                  # search variables
grep('singlefav', names(selected_nzes2011), value = TRUE) # search singlefav variable

# count vote
selected_nzes2011 %>%
  group_by(jpartyvote) %>%
  summarise(count = n())                                  # `"Don't know"` and `NA`.

# only certainty people
selected_nzes2011 %>%
  filter(jpartyvote != "Don't know") %>%
  group_by(jpartyvote) %>%
  summarise(count = n())


# count favourite
selected_nzes2011 %>%
  group_by(X_singlefav) %>%
  summarise(count = n())

# count known favourite
selected_nzes2011 %>%
  filter(!is.na(X_singlefav)) %>%  # ! : not
  group_by(X_singlefav) %>%        
  summarise(count = n())


# filter on multiple variables
selected_nzes2011 %>%
  filter(!is.na(X_singlefav), jpartyvote != "Don't know") %>%
  group_by(X_singlefav) %>%
  summarise(count = n())


# make sameparty
selected_nzes2011 <- selected_nzes2011 %>%
  mutate(sameparty = ifelse(jpartyvote == X_singlefav, 'same', 'different'))

# view sameparty
selected_nzes2011 %>%
  group_by(jpartyvote, X_singlefav, sameparty) %>%
  summarise(count = n())

# filter keep same
selected_nzes2011 %>%
  group_by(jpartyvote, X_singlefav, sameparty) %>%
  summarise(count = n()) %>%
  filter(sameparty == 'same')

# filter keep different
selected_nzes2011 %>%
  group_by(jpartyvote, X_singlefav, sameparty) %>%
  summarise(count = n()) %>%
  filter(sameparty == 'different')

# filter keep na
selected_nzes2011 %>%
  group_by(jpartyvote, X_singlefav, sameparty) %>%
  summarise(count = n()) %>%
  filter(is.na(sameparty))
```


## Q2. Relationship between age of voters and how much they like the NZ First party

```{r}
str(selected_nzes2011$jnzflike)              # factor
str(selected_nzes2011$jage)                  # numeric

# numeric summary
selected_nzes2011 %>%
  summarise(agemean = mean(jage), agemedian = median(jage), agesd = sd(jage),
            agemin = min(jage), agemax = max(jage))   #cuz, not reporting their age

# mumeric summary no na
selected_nzes2011 %>%
  filter(!(is.na(jage))) %>%
  summarise(agemean = mean(jage), agemedian = median(jage), agesd = sd(jage),
            agemin = min(jage), agemax = max(jage))
```


### Approach 1 : Strongly liking and disliking NZ First and age
# %in% : filtering for multiple values

```{r}
# pick from list
selected_nzes2011 %>%
  filter(jnzflike %in% c('0','10')) %>%
  group_by(jnzflike) %>%
  summarise(count = n())
```


### Approach 2 : Age and liking for NZ First

```{r}
# groups from threshold(numeric variable -> categorical variable)
selected_nzes2011 <- selected_nzes2011 %>%
  mutate(retiredage = ifelse(jage >= 65, 'retired age', 'working age'))
selected_nzes2011 %>%
  group_by(retiredage) %>%        
  summarise(count = n())


# numeric data error
selected_nzes2011 %>%
  group_by(retiredage) %>%
  summarise(medlike = median(jnzflike))

# change the type of the function, as.thingtochangeto(), error-version
selected_nzes2011 <- selected_nzes2011 %>%
  mutate(numlikenzf = as.numeric(jnzflike))

# check it work right way
selected_nzes2011 %>%
  group_by(jnzflike, numlikenzf) %>%
  summarise(count = n())


# change the type of the function, correct conversion
selected_nzes2011 <- selected_nzes2011 %>%
  mutate(numlikenzf = as.numeric((as.character(jnzflike))))

# checking correct conversion
selected_nzes2011 %>%
  group_by(jnzflike, numlikenzf) %>%
  summarise(count = n())
```