---
title: "Introduction to Probaility"
Week 3: 'Probaility'
---

### Getting Started

### Load package

```{r}
library(statsr)
library(dplyr)
library(ggplot2)
```


### Data

```{r}
data("kobe_basket")

names(kobe_basket)  
?kobe_basket
str(kobe_basket)    # 133 obs, 6 variables
```


### Analysis

## Q1. A streak length of 1 means one ___ followed by one miss.\[ \textrm{H M | M | H H M | M | M | M} \] => H - hit, M - miss. Their lengths are one, zero, two, zero, zero, zero 


## Q2. A streak length of 0 means one ___ which must occur after a miss that ended the preceding streak.


## Q3. False about the distribution of Kobe¡¯s streak lengths from the 2009 NBA finals.

```{r}
# counting streak lengths(variable)
kobe_streak <- calc_streak((kobe_basket$shot))
# plot - strak-kobe
ggplot(data = kobe_streak, aes(x = length)) +
  geom_histogram(binwidth = 1)

```


### Simulation in R

```{r}
# simulate flipping a fair coin
coin_outcomes <- c("heads", "tails")
sample(coin_outcomes, size = 1, replace = TRUE)

# fair coin
sim_fair_coin <- sample(coin_outcomes, size = 100, replace = TRUE)
sim_fair_coin
table(sim_fair_coin)

# unfair coin
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE, 
                          prob = c(0.2, 0.8))
sim_unfair_coin
table(sim_unfair_coin)
```


### Simulating the Independent Shooter

```{r}
shot_outcomes <- c('H', 'M')
sim_basket <- sample(shot_outcomes, size = 100, replace = TRUE)

sim_streak <- calc_streak(sim_basket)
ggplot(data = sim_streak, aes(x = length))+
  geom_histogram(binwidth = 1)

```


## Q4. If you were to run the simulation of the independent shooter a second time, how would you expect its streak distribution to compare to the distribution from the exercise above?


## Q5. How does Kobe Bryant¡¯s distribution of streak lengths compare to the distribution of streak lengths for the simulated shooter? Using this comparison, do you have evidence that the hot hand model fits Kobe¡¯s shooting patterns?